---
title: "P-values post Selection by Aggregate Testing (PSAT)"
author: "Ruth Heller"
date: "2017 - 02 - 23"
output:
  html_document: default
  pdf_document: default
vignette: |
  %\VignetteIndexEntry{Vignette Title} %\VignetteEngine{knitr::rmarkdown} %\VignetteEncoding{UTF-8}
---

#Overview#
PSAT is an R package for computing conditional p-values following selection with an aggregate level test statistic. One of the main function, aggregatePvalues, computes conditional p-values after selection based on global null p-values. Currently, selection by Fisher and Pearson global null p-values are implemented. See details in our [paper](http://dx.doi.org/10.1101/058404).

#Examples#
The following are examples of computing conditional p-values, conditional on the fact that the global null p-value was below a predefined selection threshold. 

First, we generate 2 features (rows), each examined in 20 studies (columns). We compute the one-sided and two-sided individual level p-values for these featrures. 
```{r, eval = TRUE,echo=T}
library(PSAT)

set.seed(123)
  beta  = c(rnorm(10,0,2.5), rep(0, 32))

  pmat1sided <- matrix(1-pnorm(rnorm(40,beta)), nrow=2, ncol=20, byrow=TRUE) #one-sided p-values
  print(round(pmat1sided,3))
  
  pmat2sided = 2*pmin(pmat1sided, 1-pmat1sided) #two-sided p-values
  print(round(pmat2sided,3))  

```

The one-sided Figher global null p-value and conditional p-values given that Fisher's global null p-value is at most 0.001 are computed using the function `aggregatePvalues` as follows. Note that if a feature (row) has global null p-value above the selection threshold, the conditional p-values are set as NA's. 
```{r,eval = TRUE, fig.width=6, fig.align='center',echo=T}

  out.Fisher= aggregatePvalues(pmat1sided, globaltest = "Fisher", pthreshold=0.001)
  print(out.Fisher)
  
  #plot the original p-values, and the conditional p-values after global null selection by Fisher, 
  # for the selected row. 
  par(mfrow=c(1,1))
  plot(seq(1:20), -log10(pmat1sided[1,]), col="red", ylab = "-log10(PV)", 
     main="conditional and original one-sided p-values on -log10 scale" )
  points(seq(1:20), -log10(out.Fisher$p2C[1,]), pch = 2)
  legend(12,8, c("original",  "cond Fisher"), pch = c(1,2), col=c("red", "black"))
  abline(-log10(0.05/20),0,lty=2, col="gray")



```



The two-sided Fisher global null p-value and conditional p-values given that two-sided Fisher's global null p-value is at most 0.001 are computed using the function `aggregatePvalues` as follows. 
```{r,eval = TRUE, fig.width=6, fig.align='center',echo=T}

  out.Fisher= aggregatePvalues(pmat2sided, globaltest = "Fisher", pthreshold=0.001)
  print(out.Fisher)
  
  #plot the original p-values, and the conditional p-values after global null selection by Fisher,
  # for the selected row. 
  par(mfrow=c(1,1))
  plot(seq(1:20), -log10(pmat2sided[1,]), col="red", ylab = "-log10(PV)", 
    main="conditional and original two-sided p-values on -log10 scale" )
  points(seq(1:20), -log10(out.Fisher$p2C[1,]), pch = 2)
  legend(12,8, c("original",  "cond Fisher"), pch = c(1,2), col=c("red", "black"))
  abline(-log10(0.05/20),0,lty=2, col="gray")



```


The Pearson global null p-value and conditional p-values given that Pearson's global null p-value is at most 0.001 are computed using the function `aggregatePvalues` as follows. 
```{r,eval = TRUE, fig.width=6, fig.align='center',echo=T}

  out.Pearson = aggregatePvalues(pmat1sided, globaltest = "Pearson", pthreshold=0.001)
  print(out.Pearson)

  #plot the original p-values, and the conditional p-values after global null selection by Pearson, 
  # for the selected row. 
  par(mfrow=c(1,1))
  plot(seq(1:20), -log10(pmat2sided[1,]), col="red", ylab = "-log10(PV)",
    main="conditional and original two-sided p-values on -log10 scale" )
  points(seq(1:20), -log10(out.Pearson$p2C[1,]), pch = 2)
  legend(12,8, c("original",  "cond Pearson"), pch = c(1,2), col=c("red", "black"))
  abline(-log10(0.05/20),0,lty=2, col="gray")


```



#Guidelines on when to use  2-sided Fisher versus Pearson global null p-values 

For selection of features (i.e., the rows), we combined p-values across the different  (columns). If the individual level alternatives are two-sided, we can do it either by using Fisher's combining function on two-sided p-values are  by  a  Fisher style test suggested by Pearson,  which  runs the Fisher combining method for left-sided alternatives, and separately for right-sided alternatives, and takes the maximum of the two resulting statistics (see Section 5 of our [paper](http://dx.doi.org/10.1101/058404) for details).  This test has a strong preference for common directionality (i.e., it will have greater power than a test based on Fisher's combining method on two-sided p-values when the direction of the signal is consistent across columns), while not requiring us to know the common direction.

In the example above, there was no favored direction in our simulated signal and the two-sided Fisher global null p-values was smaller than Pearson's global null p-value for the first feature (and the conditional p-values following selection by the two-sided Fisher global null p-value were smaller than the conditional p-values following selection by Pearson's global null p-value). However, in many applications, e.g., the cross-tissue eQTL analysis in The Cancer Genome Atlas project (described in Section 5 of our [paper](http://dx.doi.org/10.1101/058404)), if there is an association it is more likely in the same direction in most studies (columns), and hence Pearson's global null p-values will tend to be smaller than the two-sided Fisher global null p-values. In such case, Pearson's global null p-values should be used for greater power of the post-selection inference based on the conditional p-values.  

